---
title: "W2"
author: Liying Wang
output: html_document
---

My potential questions for this quantitative research project as following:

1. What is the trend of research of social inequality in archaeology in the past decades? Is it popular in any particular time period?

2. How often social inequality relates to culture contact? Especially for European contact. Also, does any area show preference of such research, such as North America, Southeast Asia, Oceania, Africa.    

3. For the method, what is the relationship between the term "inequality" and other archaeological evidence, including burials, ceramics, post holes, or glass beads.

In order to run the codes for my questions, I will use following two libraries: 

```{r data_input}
library(JSTORr)
# multiple_archives <- JSTOR_unpack_multiple_archives(mydir = "data/one_gram")
# save(multiple_archives, file= "data/one_gram/multiple_archives.RData" )
load("data/one_gram/multiple_archives.RData")
```

For the first question, the plots show that the topic of inequality becomes common after maybe 1992, and has became popular around 2006. The term "social is commener than the term "inequality". I think the term, "social", is too general, which make such difference. Third graph shows that social and inequality were dicussesed together after 1978, but slightly goes down at 1986, and then popular again from 1994.    
```{r 1gram}
# one word over time
JSTOR_1word(multiple_archives, "inequality")
# two words over time
JSTOR_2words(multiple_archives, "social", "inequality")
# correlation of words over time
JSTOR_2wordcor(multiple_archives, "social", "inequality")
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r 1gram}
# one word over time
JSTOR_1word(multiple_archives, "inequality")
# two words over time
JSTOR_2words(multiple_archives, "social", "inequality")
# correlation of words over time
JSTOR_2wordcor(multiple_archives, "social", "inequality")
```

```{r 1gram}
# subset samples
nouns<- JSTOR_dtmofnouns(multiple_archives, word = NULL, sparse =1, POStag = TRUE)
# Principal component analysis for articles containing inequality
PCA_inequality <- JSTOR_clusterbywords(nouns, "inequality", f = 0.005)
# Top words over time
ineq_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'inequality', sparse =1, POStag = TRUE)
ineq_freqwords <- JSTOR_freqwords(multiple_archives, ineq_nouns, n =3 )
```

