---
author: Liying Wang
title: " Examining the topic of social inequality in archaeology: An quantitative approach"
output: html_document
bibliography: myfile.bib
---

### Introduction

The development of human societies is one of the important topic in archaeology. For the studies of complex societies, social inequality is both a model and mechanism for understanding the process. I will examine how archaeologist used the model of social inequality to explain prehistoric societies. Social inequality refers to power relation and principle of hierarchical structure in human society, and usually associated with social complexity [Ames2007][@Bowles2010][@Price2010]. Social complexity is a broader topic for research the development of social organization, and refers to a consequence of evolutionary process of growing differentiation and maintenance of the structural asymmetry [@Laguens2014][Arponen2015]. Although these two terms are similar, social inequality provides a more anthropological way to examine the mechanism of complex societies compared to social complexity. The development of social inequality and social complexity can reflect different theoretical approaches to archaeology. In addition, social inequality is usually associated with some causes, for example, culture contact. I am also interested in how often do archaeologists discuss inequality and culture contact together. Besides, culture contact as a cause for social inequality might shows different patterns of popularity both in temporal and spatial scale.  

In order to figure out the development of social inequality and associated topics, I will focus on three aspects: the development of the explanatory model of social inequality, the relationship between social inequality and culture contact, and the archaeological evidence relates to this model. I will explore the model of explanation of "social inequality" in archaeology by using R as a quantitative method, and investigate how is the development of this model based on a large sample of archaeological literature over 50 years through JSTOR DfR dataset. My research questions for this project as following:   

1. What is the trend of the explanatory model of social inequality in archaeology in the past decades? Is it popular at any particular time? How is the relationship between social inequality and social complexity?

2. How often social inequality relates to culture contact? Especially for European contact.  Also, does any area show preference of such research, such as America, Asia, Oceania, or Africa?    

3. For the interpretation, what is the relationship between the term "inequality" and other archaeological evidence, including burials, ceramics, household, and glass beads.

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, cache=TRUE, warning=FALSE)
```

```{r data_input}
library(JSTORr)
# multiple_archives <- JSTOR_unpack_multiple_archives(mydir = "data/one_gram")
# save(multiple_archives, file= "data/one_gram/multiple_archives.RData" 
load("data/one_gram/multiple_archives.RData")
# library(devtools)
# download and install the package (do this only once ever per computer)
# install_github("benmarwick/JSTORr")
```

```{r load_data_bigram}
# setwd("data/two_gram")
# bigrams <- JSTOR_unpack2grams()
# save(bigrams, file = "bigrams.RData")
load("data/two_gram/bigrams.RData")
```

```{r subset_data, results='hide'}
nouns_Asia <- JSTOR_dtmofnouns(multiple_archives, word = "asia", sparse =1, POStag = TRUE)
nouns_Ame <- JSTOR_dtmofnouns(multiple_archives, word = "america", sparse =1, POStag = TRUE)
nouns_Oce <- JSTOR_dtmofnouns(multiple_archives, word = "oceania", sparse =1, POStag = TRUE)
nouns_Afr <- JSTOR_dtmofnouns(multiple_archives, word = "africa", sparse =1, POStag = TRUE)
```


### Data Analysis

####Frequencies and correlation of “social” and “inequality”
For the first question of the trend of social inequality, I examine the frequency of the word ”inequality” over time, frequencies of two words “social” and “inequality”, and the correlation between these two words.  The plots show that the topic of inequality becomes common after maybe 1992, and then has became popular around 2006. Compared to inequality, the term "social" is commoner than the term "inequality". I think the reason is that this term "social" is too general. The third plot in this set shows that social and inequality were discussed together after 1978, and there are a couple of articles show strong correlation at 1978, 1985, and 2007. Based on the results, I further explore the exact articles with strong correlation between “social” and “inequality”.  These article shows that the discussion of “social” and “inequality” shifts from exploring the causes to emphasis of cultural diversity and different type of inequality. That is, recent article suggests that there is no clear line between equality and inequality, and both principles can exist in a society [@Crumley1995]. This change also indicates different theoretical frameworks in archaeology, explanatory approach and interpretative approach [@Fogelin2007][@Jones2013]. The following are briefly annotated bibliography with strong correlation:

1. Sandra Montón Subías, 2007, Interpreting Archaeological Continuities: An Approach to Transversal Equality in the Argaric Bronze Age of South-East Iberia. World Archaeology, Vol. 39, No. 2, The Archaeology of Equality, pp. 246-262. Taylor & Francis, Ltd 

In this paper, the author explored the social inequality based on women’s material culture and concluded that the concept of social transversal equality might be more appropriate to explain the discontinuities in material culture. 

2. Bender, Barbara. "Emergent tribal formations in the American midcontinent."American Antiquity (1985): 52-62.

This author explored the cause for increasingly social differentiation in gather-hunter societies, and suggested that the alliance and exchange might exacerbate the social differentiation and inequality. 

3. Bender, Barbara. "Gatherer‐hunter to farmer: A social perspective." World archaeology 10.2 (1978): 204-222.

This paper focused on some properties of social organization among gatherer-hunters, and suggested that the development of social relations may promote the economic change.  

```{r onegram}
# one word over time
inequality <- JSTOR_1word(multiple_archives, "inequality")
inequality$plot
inequality_tbl <- inequality$word_by_year
# inequality_tbl[which.max(inequality_tbl$word_by_year), ]
# two words over time
two_soc_ine <- JSTOR_2words(multiple_archives, "social", "inequality")
two_soc_ine$plot
two_soc_ine_tbl <- two_soc_ine$words_by_year
# two_soc_ine_tbl[which.max(two_soc_ine_tbl$value), ]
# correlation of words over time
cor_soc_ine <- JSTOR_2wordcor(multiple_archives, "social", "inequality")
cor_soc_ine$plot
soc_ine_paper <- two_soc_ine_tbl[(two_soc_ine_tbl$year == 1978), ]
# soc_ine_paper[which.max(soc_ine_paper$value), ]
```

Figure1: Frequencies of word “inequality” and “social” and their correlation


####Frequencies and correlation of “social inequality” and “social complexity”
If I use "social inequality" as a term, the plots show that such discussion only exists after 1978, becomes popular around 2000. In archaeology, social inequality is usually associated with social complexity. Therefore, I examine the discussion of social complexity in order to have a general understanding of those explanatory models and their relationship. The discussion of "social complexity" began from 1947, became popular around 1978, slightly declined at 1990 and 2006, and then was popular again at 2010. When we examine the frequencies of these two terms at a same plot, it shows that they have different patterns of frequency. When social inequality was frequently discussed, the frequency of social complexity became lower, and vice versa. This pattern could be observed at 1991, 1995, 2000, and 2005, which indicates that the terms of social inequality and social complexity might have some relationship of mutually exclusive. The results suggest that the discussion of social inequality could be viewed as more specific approach to complex societies, whereas social complexity is a broader topic which covers different discussion of social expects. Moreover, compared to social complexity, the discussion of social inequality emerged later around 1978, which implied that archaeologist focus more on social relation and power, which shows the emphasis of anthropological thinking. This might relates to the development of post-processual archaeology from 1980 [@Shanks2007]. Besides, I also examine the correlation between social inequality and social complexity. The plot shows that they are strongly correlated at 1987, 1998, and 2007, which shows the periodic pattern of discussion, and might relate to review articles.   

```{r bigrams1}
# one word over time
social_inequality <- JSTOR_1bigram(bigrams, "social inequality")
social_inequality
social_inequality_tbl <- social_inequality$data
# social_inequality_tbl[which.max(social_inequality_tbl$bigram_ratio), ]
bi_socialcomplexity <- JSTOR_1bigram(bigrams, "social complexity")
bi_socialcomplexity
# correlation of words over time
bi_ineq_comp <- JSTOR_2bigrams(bigrams, "social inequality", "social complexity")
bi_ineq_comp
bicor_comp_ineq <- JSTOR_2bigramscor(bigrams, "social inequality", "social complexity")
bicor_comp_ineq
```

Figure2: Frequencies of word “social inequality” and “social complexity” and their correlation

####Correlation of “social inequality” and “culture contact”
For the second question about the correlation between social inequality and culture contact, the plot of frequencies over time shows that there was a similar trend before 1994, but after 1994 there was no clear pattern shows the relationship. When we further examine the correlation between these two terms, the plot indicates that there are a couple of articles with strong correlation at 1986, 1994, and 2007. If we limit the culture contact to only European contact, it shows that frequencies have similar pattern as the plot of culture contact. However, there is only one article with strong correlation between social inequality and European contact at 1986. This indicates that European contact might be not a common discussion for the topic of social inequality, or sometimes archaeologists use the term culture contact to incorporate the term European contact. 

```{r bigrams}
# two words over time
bi_ineq_cul <- JSTOR_2bigrams(bigrams, "social inequality", "culture contact")
bi_ineq_cul
bi_ineq_euro <- JSTOR_2bigrams(bigrams, "social inequality", "european contact")
bi_ineq_euro
# correlation of words over time
bicor_ineq_cul <- JSTOR_2bigramscor(bigrams, "social inequality", "culture contact")
bicor_ineq_cul
bicor_ineq_euro <- JSTOR_2bigramscor(bigrams, "social inequality", "european contact")
bicor_ineq_euro
```

Figure3: Frequencies and correlation between “social inequality” and “culture contact”

####Frequencies of “inequality” at different area
I further examine the topic of inequality in terms of different areas, including Asia, Oceania, America, and Africa. The results show that in Asia, most articles contain inequality exist after 1978, and the term reached the highest frequency at 2004, but declined later at 2009. For Oceania, the trend shows there are two periods with higher frequency, which are 1996 and 2004. For America, there are more articles refer to inequality, and this term was popular around 2005. The pattern of frequency in Africa looks similar to the pattern in America, and also shows the highest frequency at 2005. Although these four plots have different trends of the frequency of inequality, it seems social inequality was popular between 2004 and 2005 among these different areas.  This indicates there might be an article covers all of these areas.  

```{r subset_data2, results='hide'}
# subset the data to different area for PCA
Asi_data <- JSTOR_subset1grams(multiple_archives,"asia")
Oce_data <- JSTOR_subset1grams(multiple_archives,"oceania")
Ame_data <- JSTOR_subset1grams(multiple_archives,"america")
Afr_data <- JSTOR_subset1grams(multiple_archives,"africa")
```

```{r onegram1}
# one word, inequality, in different area
ineq_Asi <- JSTOR_1word(Asi_data, "inequality")
ineq_Asi$plot
ineq_Oce <- JSTOR_1word(Oce_data, "inequality")
ineq_Oce$plot
ineq_Ame <- JSTOR_1word(Ame_data, "inequality")
ineq_Ame$plot
ineq_Afr <- JSTOR_1word(Afr_data, "inequality")
ineq_Afr$plot
```

Figure4: Frequencies of “inequality” in Asia, Oceania, America, and Africa

#### Archaeological evidence relates to social inequality
For the third questions, these plots show the relationship between the term 'inequality' and other archaeological evidence, including beads, burials, and pottery. For the beads, it indicates that a couple of articles have strong correlation around 1990. In addition, there is no much discussion until 1976. For the burials, it also shows that the discussion of burial and inequality become common after the publication of the article in 1976. Besides, is seems the articles could be divided into two groups according to the extent of correlation. For the pottery, there is no clear correlation until 2006. After 2006, some articles show the strong correlation. The plot for house shows that there is a strong correlation at 2012. These patterns show that burials and houses are common evidence to discuss social inequality from early period to most recent period, and pottery also becomes common over time. 

```{r onegram2}
# correlation of inequality and evidence over time
cor_ineq_buri <- JSTOR_2wordcor(multiple_archives, "burial", "inequality")
cor_ineq_buri$plot
cor_ineq_pot <- JSTOR_2wordcor(multiple_archives, "pottery", "inequality")
cor_ineq_pot$plot
cor_ineq_bead <- JSTOR_2wordcor(multiple_archives, "bead", "inequality")
cor_ineq_bead$plot
cor_ineq_house <- JSTOR_2wordcor(multiple_archives, "house", "inequality")
cor_ineq_house$plot
```

Figure5: The correlation between social inequality and archaeological evidence

#### Top words related to inequality over time
The top words in all the documents containing “inequality” show the focus changes with time. The top words shift from population, organization, and mound to ritual, power and pottery. This indicates that archaeologists focus more on anthropological questions recently since 2006.  Besides, the term “settlement” could be observed in each period from 1996 until now, and it becomes more popular over time. This shows settlement is a explanatory scale in such researches.    
 
```{r onegram4, eval= TRUE, results= 'hide'}
ineq_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'inequality', sparse =1, POStag = TRUE)
# com_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'complexity', sparse =1, POStag = TRUE)
```

```{r onegram5, eval= TRUE, results= 'hide'}
# Top words containing 'inequality' over time
# exclude some irrelevant words
custom_stopwords <- c('archaeology', 'university', 'research', 'evidence', 'journal', 'world', 'site', 'cambridge', 'archaeol', 'area', 'region', 'period', 'analysis', 'anthropology', 'springer', 'production', 'figure', 'work', 'world', 'concept', 'human', 'middle', 'altamira', 'culture', 'record', 'citation', 'discipline', 'author', 'proportion', 'literature', 'report', 'approach', 'cambridge' )
ineq_freqwords <- JSTOR_freqwords(multiple_archives, ineq_nouns, custom_stopwords, n = 5)
# com_freqwords <- JSTOR_freqwords(multiple_archives, com_nouns, custom_stopwords, n = 10)
```

Figure6: The top words associated with inequality over time

###Conclusion
Based on above plots, the model of social inequality has developed after 1978, which reflects a more anthropological thinking and shows the influence of post-processual archaeology, which stressed the humanistic interpretation and diversity of approaches to explanation. The development of social inequality and social complexity represents some kind of mutually exclusive relationship.  This supports that social complexity is a broader issue, while social inequality is a specific model and also serves as a new trend for studying complex societies. Recently, ritual, or power is the common subject associated with social inequality and sheds light on future research. In addition, for the archaeological remains, houses and burials are common evidence to examine social inequality. Pottery also becomes more important in recent years. In conclusion, these results suggest that the anthropological thinking might still be the trend for future studies about complex societies, and we can focus on power or ritual based on burials and space distribution about houses to examine social inequality. 

### Reference

```{r onegram3, eval= FALSE, results='hide'}
# subset samples
nouns <- JSTOR_dtmofnouns(multiple_archives, word = NULL, sparse =1, POStag = TRUE)
```

```{r various_pca, eval= FALSE}
# Principal component analysis for articles containing inequality
PCA_inequality <- JSTOR_clusterbywords(nouns, "inequality", f = 0.005)
PCA_inequality_Asia <- JSTOR_clusterbywords(nouns_Asia, "inequality", f = 0.005)
PCA_inequality_Ame <- JSTOR_clusterbywords(nouns_Ame, "inequality", f = 0.005)
PCA_inequality_Oce <- JSTOR_clusterbywords(nouns_Oce, "inequality", f = 0.005)
PCA_inequality_Afr <- JSTOR_clusterbywords(nouns_Afr, "inequality", f = 0.005)
```

```{r various_pca2, eval= FALSE}
PCA_inequality$p
PCA_inequality$q
PCA_inequality_Asia$P
PCA_inequality_Asia$q
PCA_inequality_Ame$P
PCA_inequality_Ame$q
PCA_inequality_Oce$P
PCA_inequality_Oce$q
```
