---
title: "W6"
author: Liying Wang
output: html_document
---

My potential questions for this quantitative research project as following:

1. What is the trend of research of social inequality in archaeology in the past decades? Is it popular in any particular time period?

2. How often social inequality relates to culture contact? Especially for European contact. Also, does any area show preference of such research, such as North America, Southeast Asia, Oceania, Africa.    

3. For the method, what is the relationship between the term "inequality" and other archaeological evidence, including burials, ceramics, post holes, or glass beads.

In order to run the codes for my questions, I will use following two libraries: 

```{r data_input}
library(JSTORr)
multiple_archives <- JSTOR_unpack_multiple_archives(mydir = "data/one_gram")
save(multiple_archives, file= "data/one_gram/multiple_archives.RData" )
multiple_archives2 <- JSTOR_unpack_multiple_archives(mydir = "data/two_gram")
save(multiple_archives2, file= "data/two_gram/multiple_archives2.RData" )
load("data/one_gram/multiple_archives.RData")
load("data/two_gram/multiple_archives2.RData")
# library(devtools)
# download and install the package (do this only once ever per computer)
# install_github("benmarwick/JSTORr")
```

For the first question, the plots show that the topic of inequality becomes common after maybe 1992, and has became popular around 2006. The term "social" is commener than the term "inequality". I think the term, "social", is too general, which make such difference. Third graph shows that social and inequality were dicussesed together after 1978, but slightly goes down at 1986, and then popular again from 1994.    
```{r 1gram}
# one word over time
JSTOR_1word(multiple_archives, "inequality")
# two words over time
JSTOR_2words(multiple_archives, "social", "inequality")
# correlation of words over time
JSTOR_2wordcor(multiple_archives, "social", "inequality")
```


```{r 2gram}
# one word over time
JSTOR_1word(multiple_archives2, "social inequality")
# two words over time
JSTOR_2words(multiple_archives2, "social inequality", "European contact")
# correlation of words over time
JSTOR_2wordcor(multiple_archives2, "social inequality", "European contact")
```

The frequency of words containing “inequality” shows that the focus changes with time. The focus of paper shift from population, organization, and mound to ritual, power and pottery, which indicates there is a trend of anthropological thinking from 2006. Besdies, The term settlement could be observed in each period from 1996 until now, and it becomes more popular over time. 

```{r 1gram}
# subset samples
nouns<- JSTOR_dtmofnouns(multiple_archives, word = NULL, sparse =1, POStag = TRUE)
# Principal component analysis for articles containing inequality
PCA_inequality <- JSTOR_clusterbywords(nouns, "inequality", f = 0.005)
# Two words
PCA_social_ineq <- JSTOR_clusterbywords(nouns, c("inequality", "social"))
# Top words containing 'inequality' over time
# exclude some irrelevant words
custom_stopwords <- c('archaeology', 'university', 'research', 'evidence', 'journal', 'world', 'site', 'cambridge', 'archaeol', 'area', 'region', 'period', 'analysis', 'anthropology', 'springer', 'production', 'figure', 'work', 'world', 'concept', 'human', 'middle', 'altamira', 'culture', 'record', 'citation', 'discipline', 'author', 'proportion', 'literature', 'report', 'approach', 'cambridge' )
ineq_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'inequality', sparse =1, POStag = TRUE)
ineq_freqwords <- JSTOR_freqwords(multiple_archives, ineq_nouns, custom_stopwords, n = 5)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
