---
author: Liying Wang
title: " Examining the topic of social inequality in archaeology: An quantitative approach"
output: html_document
bibliography: myfile.bib
---

### Introduction

The development of human societies is one of important topics in archaeology. Social inequality serves as a model to understand the mechanism of complex societies. In this paper, I will explore how archaeologist used the model of social inequality to explain prehistoric societies in past years by quantitative methods. My focuses include the development of social complexity over time, causes for social complexity, and the relative archaeological evidence relates to this topic. I will conclude that social inequality serves as a new trend for studying complex societies and "power" or "ritual" is the key word to understand social inequality. 

### Hypothesis testing

My research problems is how does social inequality develop and what is the relevant topic and evidence when we talk about social inequality. The terms "social inequality" and "social complexity" were usually discussed together, and sometimes there is no clear line between them. Thus, I want to explore the relationship between social inequality and social complexity. This will be helpful to understand the topic of social inequality in studies of complex society. My hypothesis is that if the social inequality is a new trend for study complex society, then results will show the stronger relationship between social inequality and complex society in recent years compared to the relationship between social complexity and complex society. In addition, social inequality is usually associated with some causes. I would like to examine the relationship between social inequality and culture contact, especially for the expansion of European in the 17th century. If culture contact is a cause relative to social inequality, then the result will show strong correlation.   

I will explore "social inequality" in archaeology by using R as a quantitative method, and investigate how is the development of this model based on a large sample of archaeological literature over 50 years through JSTOR dataset. My research questions for this project as following:   

1. How does the development of social inequality in archaeology in past decades? Is it popular at any particular time? How is the relationship between social inequality and social complexity?

2. How often social inequality relates to culture contact? Especially for European contact.  Also, does any area show preference of such research, such as America, Asia, Oceania, or Africa?    

3. What is the relationship between the term "inequality" and other archaeological evidence, including burials, ceramics, household, and glass beads? What other evidence or conditions relates to social inequality?

### Key words

Key words will be explored in this paper include social inequality, social complexity, complex society, culture contact, European contact, and other terms relates to archaeological evidence. Based on results from above words, I will also further examine relative terms. The term social inequality refers to power relation and principle of hierarchical structure in human society, and usually associated with social complexity [@Ames2007; @Bowles2010; @Price2010]. Social complexity is a broader topic for research the development of social organization, and refers to a consequence of evolutionary process of growing differentiation and maintenance of the structural asymmetry [@Laguens2014; @Arponen2015]. 

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, cache=TRUE, warning=FALSE)
```

```{r data_input}
library(JSTORr)
# multiple_archives <- JSTOR_unpack_multiple_archives(mydir = "data/one_gram")
# save(multiple_archives, file= "data/one_gram/multiple_archives.RData" 
load("data/one_gram/multiple_archives.RData")
# library(devtools)
# download and install the package (do this only once ever per computer)
# install_github("benmarwick/JSTORr")
```

```{r load_data_bigram}
# setwd("data/two_gram")
# bigrams <- JSTOR_unpack2grams()
# save(bigrams, file = "bigrams.RData")
load("data/two_gram/bigrams.RData")
```

```{r subset_data, results='hide'}
nouns_Asia <- JSTOR_dtmofnouns(multiple_archives, word = "asia", sparse =1, POStag = TRUE)
nouns_Ame <- JSTOR_dtmofnouns(multiple_archives, word = "america", sparse =1, POStag = TRUE)
nouns_Oce <- JSTOR_dtmofnouns(multiple_archives, word = "oceania", sparse =1, POStag = TRUE)
nouns_Afr <- JSTOR_dtmofnouns(multiple_archives, word = "africa", sparse =1, POStag = TRUE)
```

### Data Analysis

####Frequencies and correlation of “social inequality” and “social complexity”

For the first question of the trend of social inequality, I examine the frequency of words of  ”social inequality” and "social complexity" over time. Figure 1 shows that such discussion only exists after 1978, becomes popular around 2000. The discussion of "social complexity" began from 1947, became popular around 1978, slightly declined at 1990 and 2006, and then was popular again at 2010. Figure 1 also shows that they have different patterns of frequency. When social inequality was frequently discussed, the frequency of social complexity became lower, and vice versa. This pattern could be observed at 1991, 1995, 2000, and 2005, which indicates that the terms of social inequality and social complexity might have some relationship of mutually exclusive. Results suggest that the discussion of social inequality could be viewed as more specific approach to complex societies, whereas social complexity is a broader topic which covers different discussion of social expects. Moreover, compared to social complexity, the discussion of social inequality emerged later around 1978, which implied that archaeologist focus more on social relation and power, which shows the emphasis of anthropological thinking. This might relates to the development of post-processual archaeology from 1980 [@Shanks2007]. 

```{r bigrams1, eval= FALSE}
# one word over time
social_inequality <- JSTOR_1bigram(bigrams, "social inequality")
social_inequality
social_inequality_tbl <- social_inequality$data
# social_inequality_tbl[which.max(social_inequality_tbl$bigram_ratio), ]
```

```{r bigrams2}
# two words over time
bi_ineq_comp <- JSTOR_2bigrams(bigrams, "social inequality", "social complexity")
bi_ineq_comp
```

Figure 1: Social inequality and social complexity over time 

I also examine the correlation between social inequality and social complexity. Figure 2 shows that they are strongly correlated at 1987, 1998, and 2007, which shows the periodic pattern of discussion, and might relate to review articles. For studies of complex society, Figure 3 and Figure 4 show that social inequality has strong correlation with complex society at 1981 and 1992, while social complexity is at 1977. This suggests that social inequality might be a trend to discuss complex society. However, compared to social complexity, there is no much correlation between social inequality and complex society after 2000.

```{r bigrams3}
# correlation of words over time
bicor_comp_ineq <- JSTOR_2bigramscor(bigrams, "social inequality", "social complexity")
bicor_comp_ineq
```

Figure 2: Correlation between social inequality and social complexity

```{r bigrams4}
# correlation of words over time
bicor_ineq_soci <- JSTOR_2bigramscor(bigrams, "social inequality", "complex society")
bicor_ineq_soci
```

Figure 3: Correlation between social inequality and complex society

```{r bigrams5}
# correlation of words over time
bicor_comp_soci <- JSTOR_2bigramscor(bigrams, "social complexity", "complex society")
bicor_comp_soci
```

Figure 4: Correlation between social complexity and complex society

In order to explore the development and key point about social inequality, I further examine articles which both contains "social" and "inequality". Fugire 5 shows that social and inequality were discussed together after 1978, and there are a couple of articles show strong correlation at 1978, 1985, and 2007. Based on results. These article shows that the discussion about social inequality shifts from exploring causes to the emphasis of cultural diversity and different type of inequality. Recent article suggests that there is no clear line between equality and inequality, and both principles can exist in a society [@Crumley1995]. This change also indicates different theoretical frameworks in archaeology, explanatory approach and interpretive approach [@Fogelin2007; @Jones2013]. Compared with explanatory approach, interpretive approach focuses more on the diversity of inequality. The following are briefly annotated bibliography with strong correlation:

1. Sandra Montón Subías, 2007, Interpreting Archaeological Continuities: An Approach to Transversal Equality in the Argaric Bronze Age of South-East Iberia. World Archaeology, Vol. 39, No. 2, The Archaeology of Equality, pp. 246-262. Taylor & Francis, Ltd 

In this paper, the author explored the social inequality based on women’s material culture and concluded that the concept of social transversal equality might be more appropriate to explain the discontinuities in material culture. 

2. Bender, Barbara, 1985, "Emergent tribal formations in the American midcontinent."American Antiquity: 52-62.

This author explored the cause for increasingly social differentiation in gather-hunter societies, and suggested that the alliance and exchange might exacerbate the social differentiation and inequality. 

3. Bender, Barbara, 1978, "Gatherer‐hunter to farmer: A social perspective." World archaeology 10.2: 204-222.

This paper focused on some properties of social organization among gatherer-hunters, and suggested that the development of social relations may promote the economic change.  

```{r onegram, eval=FALSE}
# one word over time
inequality <- JSTOR_1word(multiple_archives, "inequality")
inequality$plot
inequality_tbl <- inequality$word_by_year
# inequality_tbl[which.max(inequality_tbl$word_by_year), ]
```

```{r onegram2}
# two words over time
two_soc_ine <- JSTOR_2words(multiple_archives, "social", "inequality")
two_soc_ine_tbl <- two_soc_ine$words_by_year
# two_soc_ine_tbl[which.max(two_soc_ine_tbl$value), ]
# correlation of words over time
cor_soc_ine <- JSTOR_2wordcor(multiple_archives, "social", "inequality")
cor_soc_ine$plot
soc_ine_paper <- two_soc_ine_tbl[(two_soc_ine_tbl$year == 1978), ]
# soc_ine_paper[which.max(soc_ine_paper$value), ]
```

Figure 5: Correlation between social and inequality

####Correlation of “social inequality” and “culture contact”
For the second question about the relationship between social inequality and culture contact, Figure 6 shows that there was a similar trend before 1994, but after 1994 there was no clear pattern shows the relationship. When we further examine the correlation between these two terms, Figure 7 indicates that there are a couple of articles with strong correlation at 1986, 1994, and 2007. This means culture contact might be a possible cause for social inequality. If we limit the culture contact to only European contact, Figure 8 shows that frequencies have similar pattern as the plot of culture contact. However, there is only one article with strong correlation between social inequality and European contact at 1986. This indicates that European contact might be not a common discussion for the topic of social inequality, or sometimes archaeologists use the term culture contact to incorporate the term European contact. 

```{r bigrams6}
# two words over time
bi_ineq_cul <- JSTOR_2bigrams(bigrams, "social inequality", "culture contact")
bi_ineq_cul
```

Figure 6: Two wrods "social inequality" and "cuture contact" over time

```{r bigrams7}
# correlation of words over time
bicor_ineq_cul <- JSTOR_2bigramscor(bigrams, "social inequality", "culture contact")
bicor_ineq_cul
```

Figure 7: Correlation between social inequality and culture contact

```{r bigrams8}
bicor_ineq_euro <- JSTOR_2bigramscor(bigrams, "social inequality", "european contact")
bicor_ineq_euro
```

Figure 8: Correlation between social inequality and European contact

I further examine the topic of inequality in terms of different areas, including Asia, Oceania, America, and Africa. Results show that in Asia, most articles contain inequality exist after 1978, and the term reached the highest frequency at 2004, but declined later at 2009. For Oceania, the trend shows there are two periods with higher frequency, which are 1996 and 2004. For America, there are more articles refer to inequality, and this term was popular around 2005. The pattern of frequency in Africa looks similar to the pattern in America, and also shows the highest frequency at 2005. Although these four plots have different trends of the frequency of inequality, it seems social inequality was popular between 2004 and 2005 among these different areas.  This indicates there might be some article covers all of these areas. In addition, the frequency of inequality in America and Africa shows similar pattern, which might suggests there are some close relationship between them.     

```{r subset_data2, results='hide'}
# subset the data to different area for PCA
Asi_data <- JSTOR_subset1grams(multiple_archives,"asia")
Oce_data <- JSTOR_subset1grams(multiple_archives,"oceania")
Ame_data <- JSTOR_subset1grams(multiple_archives,"america")
Afr_data <- JSTOR_subset1grams(multiple_archives,"africa")
```

```{r onegram3_Asia}
# one word, inequality, in different area
ineq_Asi <- JSTOR_1word(Asi_data, "inequality")
ineq_Asi$plot
```

Figure 9: The word "Inequality" over time in studies relate to Asia

```{r onegram3_oceania}
ineq_Oce <- JSTOR_1word(Oce_data, "inequality")
ineq_Oce$plot
```

Figure 10: The word "Inequality" over time in studies relate to Oceania

```{r onegram3_America}
ineq_Ame <- JSTOR_1word(Ame_data, "inequality")
ineq_Ame$plot
```

Figure 11: The word "Inequality" over time in studies relate to America

```{r onegram3_Africa}
ineq_Afr <- JSTOR_1word(Afr_data, "inequality")
ineq_Afr$plot
```

Figure 12: The word "Inequality" over time in studies relate to Africa

#### Archaeological evidence relates to social inequality
For the third question, these plots show the relationship between the term 'inequality' and other archaeological evidence, including beads, burials, and pottery. For burials, it also shows that the discussion of burial and inequality become common after the publication of an article in 1976. Besides, is seems that articles could be divided into two groups according to the extent of correlation. For the pottery, there is no clear correlation until 2006. After 2006, some articles show the strong correlation. For beads, the plot indicates that a couple of articles have strong correlation around 1990. In addition, there is no much discussion until 1976. The plot for house shows that there is a strong correlation at 2012. These patterns show that burials and houses are common evidence to discuss social inequality from early period to most recent period, and pottery also becomes common over time. 

```{r onegram4_burial}
# correlation of inequality and evidence over time
cor_ineq_buri <- JSTOR_2wordcor(multiple_archives, "burial", "inequality")
cor_ineq_buri$plot
```

Figure 13: Correlation of inequality and burial

```{r onegram4_pottery}
cor_ineq_pot <- JSTOR_2wordcor(multiple_archives, "pottery", "inequality")
cor_ineq_pot$plot
```

Figure 14: Correlation of inequality and pottery

```{r onegram4_bead}
cor_ineq_bead <- JSTOR_2wordcor(multiple_archives, "bead", "inequality")
cor_ineq_bead$plot
```

Figure 15: Correlation of inequality and bead

```{r onegram4_house}
cor_ineq_house <- JSTOR_2wordcor(multiple_archives, "house", "inequality")
cor_ineq_house$plot
```

Figure 16: Correlation of inequality and house

#### Top words related to inequality over time
Top words in all documents containing “inequality” show that focuses change with time. The top words shift from population, organization, and mound to ritual, power and pottery. This indicates that archaeologists focus more on anthropological questions recently since 2006. Moreover, the term "river" and "coast" are also top words associated with inequality recently. These geographic features might indicate some conditions for social inequality. Based on results, I further explore the correlation between "inequality" and "ritual" and "power" respectively. Figure 18 shows that there are several strong correlation between ritual and inequality at 1963, and moderate correlation at 1977, 1984, 2000 and 2011. On the other hand, Figure 19 shows strong correlation between inequality and power at 1979 and 1990, and several moderate correlations from 1978 to 2011. These patterns indicate that ritual and power are common topics that are associated with inequality. In addition, the discussion of power is later than ritual.                
 
```{r onegram5, eval= TRUE, results= 'hide'}
ineq_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'inequality', sparse =1, POStag = TRUE)
```

```{r onegram6, eval= TRUE, results= 'hide'}
# Top words containing 'inequality' over time
# exclude some irrelevant words
custom_stopwords <- c('archaeology', 'university', 'research', 'evidence', 'journal', 'world', 'site', 'cambridge', 'archaeol', 'area', 'region', 'period', 'analysis', 'anthropology', 'springer', 'production', 'figure', 'work', 'world', 'concept', 'human', 'middle', 'altamira', 'culture', 'record', 'citation', 'discipline', 'author', 'proportion', 'literature', 'report', 'approach', 'cambridge', 'settlement' )
ineq_freqwords <- JSTOR_freqwords(multiple_archives, ineq_nouns, custom_stopwords, n = 5)
```

Figure17: The top words associated with inequality over time

```{r onegram7}
# two words over time
two_ri_ine <- JSTOR_2words(multiple_archives, "ritual", "inequality")
two_ri_ine
two_ri_ine_tbl <- two_ri_ine$words_by_year
#two_ri_ine_tbl[which.max(two_ri_ine_tbl$value), ]
# correlation of words over time
cor_ri_ine <- JSTOR_2wordcor(multiple_archives, "ritual", "inequality")
cor_ri_ine$plot
#ri_ine_paper <- two_ri_ine_tbl[(two_ri_ine_tbl$year == 2011), ]
#ri_ine_paper[which.max(ri_ine_paper$value), ]
```

Figure18: Correlation of inequality and ritual

```{r onegram8}
cor_pow_ine <- JSTOR_2wordcor(multiple_archives, "power", "inequality")
cor_pow_ine$plot
```

Figure19: Correlation of inequality and power

###Conclusion
Based on above plots, the model of social inequality has developed after 1978, which reflects a more anthropological thinking and shows the influence of post-processual archaeology, which stressed the humanistic interpretation and diversity of approaches to explanation. The development of social inequality and social complexity represents some kind of mutually exclusive relationship.
In addition, social inequality has strong correlation with complex society at recent years, while the strong correlation of social complexity and complex society is at eelier years. This suggests that social inequality is a specific model and might be a new trend to discuss complex society. Regarding associated topics, ritual or power is the common subject associated with social inequality in recent years and sheds light on future research. Moreover, culture contact is one of important causes relate to social inequality can be observed from the correlation plot. In addition, for archaeological remains, houses and burials are common evidence to examine social inequality. Pottery also becomes more important in recent years. In conclusion, compared with social complexity, the development of social inequality reflects an anthropological trend for studying complex society, and we can focus on power or ritual based on burials and space distribution about houses to examine social inequality. 

### Reference

```{r nouns, eval= FALSE, results='hide'}
# subset samples
nouns <- JSTOR_dtmofnouns(multiple_archives, word = NULL, sparse =1, POStag = TRUE)
```

```{r various_pca, eval=FALSE, results='hide'}
# Principal component analysis for articles containing inequality
PCA_inequality <- JSTOR_clusterbywords(nouns, "inequality", f = 0.005)
PCA_inequality_Asia <- JSTOR_clusterbywords(nouns_Asia, "inequality", f = 0.005)
PCA_inequality_Ame <- JSTOR_clusterbywords(nouns_Ame, "inequality", f = 0.005)
PCA_inequality_Oce <- JSTOR_clusterbywords(nouns_Oce, "inequality", f = 0.005)
PCA_inequality_Afr <- JSTOR_clusterbywords(nouns_Afr, "inequality", f = 0.005)
```

```{r various_pca2, eval= FALSE}
PCA_inequality$p
PCA_inequality_Asia$p
PCA_inequality_Ame$p
PCA_inequality_Oce$p
```
