---
title: "W2"
author: Liying Wang
output: html_document
---

My potential questions for this auantitative research project as following:

1. What is the trend of research of social ineaulity in archaeolgy in the past decades? Is it popular in any particular time period?

2. How often social inequality relates to culture contact? Especilally for European contact. Also, does any area show preference of such research, such as North America, Southeast Asia, Oceania, Afirca.    

3. For the method, what is the relationship between the term "inequality" and other arhcaeological evidence, including burials, ceramics, post holes, or glass beads.

In order to run the codes for my questions, I will use following two lbraries: 

```{r data_input}
library(JSTORr)
multiple_archives <- JSTOR_unpack_multiple_archives(mydir = "data")
```

For the first question, the plots show that the topic of inequality becomes common after maybe 1984.  
```{r 1gram}
# one word over time
JSTOR_1word(multiple_archives, "inequality")
# two words over time
JSTOR_2words(multiple_archives, "social", "inequality")
# correlation of words over time
JSTOR_2wordcor(multiple_archives, "social", "inequality")
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
